\LARGE \textbf{Constrained Geometric Attitude Control on \( \SO \)}

### 2017-11-25 manuscript.tex
Modify manuscript for arxiv  upload
Uses basic article class, fixes subcaption images and adds
a big header to cite the real journal paper

\documentclass{IJCAS}  % Comment this line out if you need a4paper
\usepackage{ijcas_packages}
\allowdisplaybreaks[1]

@@

%%%% Editorial Information
%% Authors do not have to modify this section.
\journalvolumn{VV}
\journalnumber{N}
\journalyear{YYYY}
\setarticlestartpagenumber{1}
%%%% End of Editorial Information

@@

\begin{document}

\title{\LARGE \textbf{Constrained Geometric Attitude Control on \( \SO \)} }

\author{Shankar Kulumani* and Taeyoung Lee}%

@@

\begin{keywords}
adaptive control, attitude control, constraint, obstacle, special orthogonal group
%At least four key words or phrases in alphabetical order, separated by commas.
\end{keywords}

\maketitle % \maketitle command must come here after ABSTRACT and KEYWORDS environments

\makeAuthorInformation{
Manuscript received DATE; revised DATE; accepted DATE.

Shankar Kulumani and Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 \texttt{ \{skulumani,tylee\}@gwu.edu}

This research has been supported in part by NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.\\
* Corresponding author.\vskip 0.5pc
}

\runningtitle{2017}{Shankar Kulumani and Taeyoung Lee}{Constrained Geometric Attitude Control on \( \SO \) }

\begin{multicols}{2}
\section{Introduction}\label{sec:intro}

Rigid body attitude control is an important problem for aerospace vehicles, ground and underwater vehicles, as well as robotic systems~\cite{hughes2004,wertz1978}.
One distinctive feature of the attitude dynamics of rigid bodies is that it evolves on a nonlinear manifold.
The three-dimensional special orthogonal group, or \( \SO \), is the set of \( 3 \times 3 \) orthogonal matrices whose determinant is one.
This configuration space is non-Euclidean and yields unique stability properties which are not observable on a linear space.
For example, it is impossible to achieve global attitude stabilization using continuous time-invariant feedback~\cite{bhat2000}.

@@

\Cref{eqn:psi} is composed of an attractive term, \( A (R) \) toward the desired attitude, and a repulsive term, \( B(R) \) away from the undesired direction \( R^T v \).
In order to visualize the attitude error function on \( \SO \), we utilize a spherical coordinate representation.
Recall, that the spherical coordinate system represents the position of a point relative to an origin in terms of a radial distance, azimuth, and elevation.
This coordinate system is commonly used to define locations on the Earth in terms of a latitude and longitude.
Similarly, the positions of celestial objects are defined on the celestial sphere in terms of right ascension and declination. 
\end{multicols}
\begin{figure}[H]%
    \centering 
    \subfigure[Attractive \( A(R) \) \label{fig:attract_error}]{\includegraphics[trim={8mm 2mm 8mm 2mm}, clip]{figures/attract_error.pdf}}~
    \subfigure[Repulsive \( B(R) \) \label{fig:avoid_error} ]{\includegraphics[trim={8mm 2mm 8mm 2mm}, clip]{figures/avoid_error.pdf}}~
    \subfigure[Combined \( \Psi \) \label{fig:combined_error}]{\includegraphics[trim={8mm 2mm 8mm 2mm}, clip]{figures/combined_error.pdf}}%
    \caption{Visualization of Configuration Error Functions using spherical coordinate representation}
    \label{fig:config_error} 
\end{figure}%
\begin{multicols}{2}
We apply this concept and parameterize the rotation matrix \( R \in \SO \) in terms of the spherical angles \( \SI{-180}{\degree} \leq \lambda \leq \SI{180}{\degree}  \) and \( \SI{-90}{\degree} \leq \beta \leq \SI{90}{\degree} \). 
Using the elementary Euler rotations, the rotation matrix is now defined as \( R = \exp( \lambda \hat{e}_2) \exp( \beta \hat{e}_3) \).
We iterate over the domains of \( \lambda\) and \(\beta\) in order to rotate the body-fixed vector \( r \) throughout the two-sphere \( \S^2 \).
Applying this method,~\cref{fig:config_error} allows us to visualize the error function on \( \SO \).
The horizontal axes of~\cref{fig:config_error} represent the domain of the spherical angles \( \lambda \) and \( \beta \) in degrees, while the vertical axes represent the unitless magnitude of the error functions defined in~\cref{eqn:psi,eqn:A,eqn:B}.
The attractive error function, given by~\cref{eqn:A}, has been previously used for attitude control on \(\SO\).
The potential well of \( A(R)\) is illustrated in~\cref{fig:attract_error}, where the desired attitude lies at the minimum of \( A(R) \).

@@

\end{multicols}
\begin{figure}[t]
    \centering 
    \subfigure[Attitude error vector \(e_R\) components \label{fig:eR_con}]{\input{pgf/eR_noadapt.pgf}}~
    \subfigure[Configuration error \( \Psi \) \label{fig:Psi_con}]{\input{pgf/Psi_noadapt.pgf}}~
    \subfigure[Angle to each constraint \label{fig:con_angles_con}]{\input{pgf/ang_con_noadapt.pgf}} 
    \caption{Attitude stabilization without adaptive update law}
    \label{fig:con} 
\end{figure}
\begin{multicols}{2}
\subsection{Attitude Control without Disturbance}
We introduce a nonlinear geometric controller for the attitude stabilization of a rigid body.
We first assume that there is no disturbance, i.e., \( \Delta = 0 \), and present a nonlinear controller in~\Cref{prop:att_control}.
\begin{prop}[Attitude Control]\label{prop:att_control}
	Given a desired attitude command \( \parenth{R_d, \Omega_d = 0} \), which satisfies the constraint~\cref{eqn:constraint}, and positive constants \( k_R, k_\Omega \in \R \) we define a control input \( u \in \R^3 \) as follows
	\begin{gather}
		u = -k_R e_R - k_\Omega e_\Omega + \Omega \times J \Omega . \label{eqn:nodist_control}
	\end{gather}
	Then the zero equilibrium of the attitude error is asymptotically stable, and the inequality constraint is always satisfied.
\end{prop}
\begin{proof}
See~\Cref{proof:att_control}
\end{proof}

@@

We demonstrate the performance of the proposed control system via numerical simulation.
The inertia tensor of a rigid body is given as
\begin{gather*}
    J = \begin{bmatrix}
	\num{5.5} & \num{0.06} & \num{-0.03} \\
	\num{0.06} & \num{5.5} & \num{0.01} \\
	\num{-0.03} & \num{0.01} & \num{0.1}
    \end{bmatrix} \times 10^{-3}~\si{\kilo\gram\meter\squared} .
\end{gather*} 
The control system parameters are chosen as
\begin{gather*}
	G = \text{diag} [0.9,1.1,1.0], \quad k_R = 0.4 , \quad	k_\Omega = 0.296 ,\\
	c = 1.0 , \quad k_\Delta = 0.5 , \quad \alpha = 15 .
\end{gather*}
The diagonal matrix \( G \) serves as a weighting matrix for the relative difference between \( R \) and \( R_d \). 
Using this term, the control designer can modify the shape of the attractive error function, given in~\cref{eqn:A}, and the resulting behavior of the closed loop system.
Similarly, the constant \( \alpha \) is used to modify the shape of the repulsive error function, given in~\cref{eqn:B}.
In general, this term is derived from the system design and the nature of the obstacles in the dynamic environment.
For example, a system wishing to avoid pointing at a diffuse obstacle, such as incoming debris, may chose an appropriate value of \( \theta \), based on the best available information, and a relatively low \( \alpha \) to ensure additional safety margin near the constraint boundary. 
Similarly, in an environment with several densely spaced obstacles, such as that presented in~\cref{fig:cad_adapt}, a much larger \( \alpha \) would enable more aggressive maneuvers which pass closer to the constraint boundary without violation.
This would increase the allowable region of operation in a highly constrained environment.
The parameters \( k_R, k_\Omega, c, k_\Delta\) are control parameters used to modify the closed-loop behavior of the system.
It is straightforward to chose \( k_R, k_\Omega, k_\Delta\), using a linear analysis, to satisfy desired response criteria, such as settling time or percent overshoot~\cite{nise2004}.
A body fixed sensor is defined as \(r = [1,0,0]\), while multiple inequality constraints are defined in~\Cref{tab:constraints}.
The simulation parameters are chosen to be similar to those found in~\cite{lee2011b}, however we increase the size of the constraint regions to create a more challenging scenario for the control system.
\vspace{-2.5mm}
\begin{table}[H]
\caption{Constraint Parameters~\label{tab:constraints}}
\begin{center}\begin{tabular}{lc}
Constraint Vector (\( v \)) & Angle (\( \theta \)) \\ \hline \hline 
\([0.174,\,-0.934,\, -0.034]^T\) & \ang{40} \\ \hline 
\([0 ,\, 0.7071 ,\, 0.7071]^T\) & \ang{40} \\ \hline 
\([-0.853 ,\, 0.436 ,\, -0.286]^T\) & \ang{40} \\ \hline 
\([-0.122 ,\,-0.140,\, -0.983]^T\) & \ang{20}\end{tabular} 
\end{center}
\end{table}
\vspace{-2.5mm}
The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\), with \( e_3 = \begin{bmatrix} 0 & 0 & 1 \end{bmatrix}^T \).
The desired state is \( R_d = I,\Omega_d = 0\).
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \si{\newton\meter}\), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e}_{R} \), is linear with respect to the angular velocity error vector \( e_\Omega\).

@@

\end{multicols}
\begin{figure} 
  \centering 
  \subfigure[Configuration error \( \Psi \)\label{fig:Psi_adapt}]{\input{pgf/Psi_adapt.pgf}}~
  \subfigure[Angle to each constraint \label{fig:con_angles}]{\input{pgf/ang_con_adapt.pgf}}\\
  \subfigure[Disturbance estimate \( \bar \Delta \) components \label{fig:delta_adapt} ]{\input{pgf/dist_adapt.pgf}}~
    \subfigure[Attitude trajectory \label{fig:cad_adapt}]{\includegraphics[trim={10cm 0 10cm 0},clip,width=0.4\columnwidth]{cad_adapt} }
  \caption{Attitude stabilization with adaptive update law}
  \label{fig:adapt} 
\end{figure}
\begin{multicols}{2}
\subsection{Attitude Parameterizations}\label{ssec:attitude_parameterization}
% Discuss the benefits of using this geometric representation as compared to Euler angles or quaternions
Attitude parameterizations, such as Euler angles and Quaternions, are frequently used in the aerospace and astrodynamics communities~\cite{vallado2007}.
For example, Euler angle sequences are frequently used to describe the transformation between a variety of reference frames used to describe the position and orientation of the orbit of Earth satellites~\cite{vallado2007}.
In addition, quaternions were used during the operation of Skylab and the NASA Space Shuttle~\cite{hughes2004}.
However, the choice of attitude parameterization plays a critical role in control design and the resulting motion of the system.

@@

\section{Experiment on Hexrotor UAV}\label{sec:experiment}
\vspace{-5mm}
\begin{figure}[H]
    \centering
    \includegraphics[width = 0.50\columnwidth]{hexrotor}
    \caption{Attitude control testbed~\label{fig:hexrotor}}
\end{figure}
\vspace{-3mm}
A hexrotor unmanned aerial vehicle (UAV), as seen in~\cref{fig:hexrotor}, has been developed at the Flight Dynamics and Controls Laboratory (FDCL) at the George Washington University~\cite{kaufman2014}.
The UAV is composed of three pairs of counter-rotating propellers. 
Typical UAVs are composed of four or more co-planar propellers.
As a result, these systems are underactuated and unable to impart a force along every degree of freedom.
For example, quadrotor UAVs are unable to translate laterally without first conducting a rotation.
Conversely, the propeller pairs of the hexrotor are angled relative to one another to allow for a fully actuated rigid body.
This allows the hexrotor to impart a force in any direction and a moment about any axis. 
\end{multicols}
\begin{figure}[H]
    \centering 
    \subfigure[{Configuration error \( \Psi \)} \label{fig:Psi_tv} ]{\input{pgf/Psi_timevarying.pgf}}~
    \subfigure[{Angle to constraint } \label{fig:con_angle_tv}]{\input{pgf/ang_con_timevarying.pgf}}~
    \subfigure[{Disturbance estimate \(\bar\Delta\) components} \label{fig:Delta_tv}]{\input{pgf/dist_timevarying.pgf}}
    \caption{Time-varying external disturbance simulation}
    \label{fig:tv} 
\end{figure}
\begin{figure}[t]
    \centering 
    \subfigure[{Attitude error vector \(e_R\)} components\label{fig:eR_exp}]{\input{pgf/eR_exp.pgf}}~
    \subfigure[{Configuration error \( \Psi \)} \label{fig:Psi_exp} ]{\input{pgf/Psi_exp.pgf}}~
    \subfigure[{Control input \( u\) } components\label{fig:u_exp}]{\input{pgf/u_exp.pgf}}\\
    \subfigure[{Attitude Trajectory } \label{fig:traj_exp}]{\includegraphics[width=0.3\columnwidth]{traj_exp} }~
    \subfigure[{Angle to constraint } \label{fig:con_angle_exp}]{\input{pgf/ang_con_exp.pgf}}
    \caption{Constrained Attitude stabilization experiment}
    \label{fig:exp} 
\end{figure}
\begin{multicols}{2}
Attitude information is measured by a combination of both on and off board sensor systems.
The VectorNav VN-100 is a rugged, miniature high-performance inertial measurement unit which provides high frequency angular velocity measurements.
A Vicon motion capture system is installed within the test environment and used to provide high accuracy attitude measurements. 
A series of reflective markers are placed on the hexrotor and their relative position is captured by a series of infrared optical cameras. 
Assuming a fixed rigid body, the Vicon system is able to derive the attitude of the hexrotor and transmit this data to the processor onboard the hexrotor.
The control input is computed on-board, using the full state measurement, and implemented at approximately \SI{100}{\hertz}.
In order to constrain the motion, allowing us to test only the attitude dynamics, we attach the hexrotor to a spherical joint.
Since, the center of rotation is below the center of gravity of the hexrotor there is a destabilizing gravitational moment.
The resulting attitude dynamics are similar to an inverted pendulum model.
We augment the control input in~\cref{eqn:adaptive_control} with an additional term to negate the effect of the gravitational moment.

@@

\bibliography{BibMaster,library}
\bibliographystyle{IEEEtran}
\biography{profile}{Shankar Kulumani}
    {is a PhD Candidate in the Department of Mechanical and Aerospace Engineering at George Washington University. 
    He received his bachelor's degree in Astronautical Engineering from the US Air Force Academy in 2009 and his master's degree in Aeronautical and Astronautical Engineering from Purdue University in 2013.
    % He previously served as Deputy Space Vehicles Lead for the ORS-1 (Operationally Responsive Space ) satellite program, and as Lead Test Engineer in the Guidance, Navigation, and Control group at Air Force Research Laboratory located at Kirtland AFB, NM.
    % He is now currently a reserve air force officer where he serves as Threat Systems Engineer for the Missle and Space Intelligence Center, Defense Intelligence Agency.
    His current research interests include the application of geometric mechanics and control to aerospace systems. 
    }
\biography{Lee}{Taeyoung Lee}
    {is an associate professor of Department of Mechanical and Aerospace Engineering at the George Washington University. 
    He received his doctoral degree in Aerospace Engineering and his master's degree in Mathematics at the University of Michigan in 2008. 
    His research interests include computational geometric mechanics and control of complex systems.}

\end{multicols}

>>>>

\documentclass[10pt]{article}  % Comment this line out if you need a4paper
\usepackage{ijcas_packages}
\usepackage{fancyhdr}
\usepackage{bibentry}
\setlength{\headheight}{15.2pt}
\pagestyle{fancyplain}
\lhead[]{}
\chead[]{}
\rhead[]{}
\fancyhf[C]{\color{red}{Please use citation: S. Kulumani and T. Lee, ``Constrained Attitude Control on \(\SO\),'' International Journal of Control, Automation, and Systems, vol. 15, no. 6, Dec. 2017}}
\allowdisplaybreaks[1]

@@

@@

\title{\LARGE \textbf{Constrained Geometric Attitude Control on \( \SO \)} }
\author{Shankar Kulumani* and Taeyoung Lee}%

\begin{document}
\maketitle

@@

\section{Introduction}\label{sec:intro}
Rigid body attitude control is an important problem for aerospace vehicles, ground and underwater vehicles, as well as robotic systems~\cite{hughes2004,wertz1978}.
One distinctive feature of the attitude dynamics of rigid bodies is that it evolves on a nonlinear manifold.
The three-dimensional special orthogonal group, or \( \SO \), is the set of \( 3 \times 3 \) orthogonal matrices whose determinant is one.
This configuration space is non-Euclidean and yields unique stability properties which are not observable on a linear space.
For example, it is impossible to achieve global attitude stabilization using continuous time-invariant feedback~\cite{bhat2000}.

@@

\Cref{eqn:psi} is composed of an attractive term, \( A (R) \) toward the desired attitude, and a repulsive term, \( B(R) \) away from the undesired direction \( R^T v \).
In order to visualize the attitude error function on \( \SO \), we utilize a spherical coordinate representation.
Recall, that the spherical coordinate system represents the position of a point relative to an origin in terms of a radial distance, azimuth, and elevation.
This coordinate system is commonly used to define locations on the Earth in terms of a latitude and longitude.
Similarly, the positions of celestial objects are defined on the celestial sphere in terms of right ascension and declination. 
\begin{figure}[htbp]%
    \centering 
    \subcaptionbox{ Attractive \( A(R) \) \label{fig:attract_error} }{\includegraphics[trim={8mm 2mm 8mm 2mm}, clip, width=0.3\textwidth]{figures/attract_error.pdf}}~
    \subcaptionbox{ Repulsive \( B(R) \) \label{fig:avoid_error} }{\includegraphics[trim={8mm 2mm 8mm 2mm}, clip, width=0.3\textwidth]{figures/avoid_error.pdf}}~
    \subcaptionbox{ Combined \( \Psi \) \label{fig:combined_error} }{\includegraphics[trim={8mm 2mm 8mm 2mm}, clip, width=0.3\textwidth]{figures/combined_error.pdf}}%
    \caption{Visualization of Configuration Error Functions using spherical coordinate representation}
    \label{fig:config_error} 
\end{figure}%
We apply this concept and parameterize the rotation matrix \( R \in \SO \) in terms of the spherical angles \( \SI{-180}{\degree} \leq \lambda \leq \SI{180}{\degree}  \) and \( \SI{-90}{\degree} \leq \beta \leq \SI{90}{\degree} \). 
Using the elementary Euler rotations, the rotation matrix is now defined as \( R = \exp( \lambda \hat{e}_2) \exp( \beta \hat{e}_3) \).
We iterate over the domains of \( \lambda\) and \(\beta\) in order to rotate the body-fixed vector \( r \) throughout the two-sphere \( \S^2 \).
Applying this method,~\cref{fig:config_error} allows us to visualize the error function on \( \SO \).
The horizontal axes of~\cref{fig:config_error} represent the domain of the spherical angles \( \lambda \) and \( \beta \) in degrees, while the vertical axes represent the unitless magnitude of the error functions defined in~\cref{eqn:psi,eqn:A,eqn:B}.
The attractive error function, given by~\cref{eqn:A}, has been previously used for attitude control on \(\SO\).
The potential well of \( A(R)\) is illustrated in~\cref{fig:attract_error}, where the desired attitude lies at the minimum of \( A(R) \).

@@

\begin{figure}[t]
    \centering 
    \subcaptionbox{Attitude error vector \(e_R\) components \label{fig:eR_con}}{\input{pgf/eR_noadapt.pgf}}~
    \subcaptionbox{Configuration error \( \Psi \) \label{fig:Psi_con}}{\input{pgf/Psi_noadapt.pgf}}~
   \subcaptionbox{Angle to each constraint \label{fig:con_angles_con}}{\input{pgf/ang_con_noadapt.pgf}}~
    \caption{Attitude stabilization without adaptive update law}
    \label{fig:con} 
\end{figure}
\subsection{Attitude Control without Disturbance}
We introduce a nonlinear geometric controller for the attitude stabilization of a rigid body.
We first assume that there is no disturbance, i.e., \( \Delta = 0 \), and present a nonlinear controller in~\Cref{prop:att_control}.
\begin{prop}[Attitude Control]\label{prop:att_control}
	Given a desired attitude command \( \parenth{R_d, \Omega_d = 0} \), which satisfies the constraint~\cref{eqn:constraint}, and positive constants \( k_R, k_\Omega \in \R \) we define a control input \( u \in \R^3 \) as follows
	\begin{gather}
		u = -k_R e_R - k_\Omega e_\Omega + \Omega \times J \Omega . \label{eqn:nodist_control}
	\end{gather}
	Then the zero equilibrium of the attitude error is asymptotically stable, and the inequality constraint is always satisfied.
\end{prop}
\begin{proof}
See~\Cref{proof:att_control}
\end{proof}

@@

We demonstrate the performance of the proposed control system via numerical simulation.
The inertia tensor of a rigid body is given as
\begin{gather*}
    J = \begin{bmatrix}
	\num{5.5} & \num{0.06} & \num{-0.03} \\
	\num{0.06} & \num{5.5} & \num{0.01} \\
	\num{-0.03} & \num{0.01} & \num{0.1}
    \end{bmatrix} \times 10^{-3}~\si{\kilo\gram\meter\squared} .
\end{gather*} 
The control system parameters are chosen as
\begin{gather*}
	G = \text{diag} [0.9,1.1,1.0], \quad k_R = 0.4 , \quad	k_\Omega = 0.296 ,\\
	c = 1.0 , \quad k_\Delta = 0.5 , \quad \alpha = 15 .
\end{gather*}
The diagonal matrix \( G \) serves as a weighting matrix for the relative difference between \( R \) and \( R_d \). 
Using this term, the control designer can modify the shape of the attractive error function, given in~\cref{eqn:A}, and the resulting behavior of the closed loop system.
Similarly, the constant \( \alpha \) is used to modify the shape of the repulsive error function, given in~\cref{eqn:B}.
In general, this term is derived from the system design and the nature of the obstacles in the dynamic environment.
For example, a system wishing to avoid pointing at a diffuse obstacle, such as incoming debris, may chose an appropriate value of \( \theta \), based on the best available information, and a relatively low \( \alpha \) to ensure additional safety margin near the constraint boundary. 
Similarly, in an environment with several densely spaced obstacles, such as that presented in~\cref{fig:cad_adapt}, a much larger \( \alpha \) would enable more aggressive maneuvers which pass closer to the constraint boundary without violation.
This would increase the allowable region of operation in a highly constrained environment.
The parameters \( k_R, k_\Omega, c, k_\Delta\) are control parameters used to modify the closed-loop behavior of the system.
It is straightforward to chose \( k_R, k_\Omega, k_\Delta\), using a linear analysis, to satisfy desired response criteria, such as settling time or percent overshoot~\cite{nise2004}.
A body fixed sensor is defined as \(r = [1,0,0]\), while multiple inequality constraints are defined in~\Cref{tab:constraints}.
The simulation parameters are chosen to be similar to those found in~\cite{lee2011b}, however we increase the size of the constraint regions to create a more challenging scenario for the control system.
\vspace{-2.5mm}
\begin{table}[htbp]
\caption{Constraint Parameters~\label{tab:constraints}}
\begin{center}\begin{tabular}{lc}
Constraint Vector (\( v \)) & Angle (\( \theta \)) \\ \hline \hline 
\([0.174,\,-0.934,\, -0.034]^T\) & \ang{40} \\ \hline 
\([0 ,\, 0.7071 ,\, 0.7071]^T\) & \ang{40} \\ \hline 
\([-0.853 ,\, 0.436 ,\, -0.286]^T\) & \ang{40} \\ \hline 
\([-0.122 ,\,-0.140,\, -0.983]^T\) & \ang{20}\end{tabular} 
\end{center}
\end{table}
\vspace{-2.5mm}
The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\), with \( e_3 = \begin{bmatrix} 0 & 0 & 1 \end{bmatrix}^T \).
The desired state is \( R_d = I,\Omega_d = 0\).
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \si{\newton\meter}\), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e}_{R} \), is linear with respect to the angular velocity error vector \( e_\Omega\).

@@

\begin{figure} 
  \centering 
  \subcaptionbox{ Configuration error \( \Psi \)\label{fig:Psi_adapt} }{\input{pgf/Psi_adapt.pgf}}~
  \subcaptionbox{ Angle to each constraint \label{fig:con_angles} }{\input{pgf/ang_con_adapt.pgf}}\\
  \subcaptionbox{ Disturbance estimate \( \bar \Delta \) components \label{fig:delta_adapt} }{\input{pgf/dist_adapt.pgf}}~
    \subcaptionbox{ Attitude trajectory \label{fig:cad_adapt} }{\includegraphics[trim={10cm 0 10cm 0},clip,width=0.4\columnwidth]{cad_adapt} }
  \caption{Attitude stabilization with adaptive update law}
  \label{fig:adapt} 
\end{figure}
\subsection{Attitude Parameterizations}\label{ssec:attitude_parameterization}
% Discuss the benefits of using this geometric representation as compared to Euler angles or quaternions
Attitude parameterizations, such as Euler angles and Quaternions, are frequently used in the aerospace and astrodynamics communities~\cite{vallado2007}.
For example, Euler angle sequences are frequently used to describe the transformation between a variety of reference frames used to describe the position and orientation of the orbit of Earth satellites~\cite{vallado2007}.
In addition, quaternions were used during the operation of Skylab and the NASA Space Shuttle~\cite{hughes2004}.
However, the choice of attitude parameterization plays a critical role in control design and the resulting motion of the system.

@@

\section{Experiment on Hexrotor UAV}\label{sec:experiment}

\begin{figure}[htbp]
    \centering
    \includegraphics[width = 0.50\columnwidth]{hexrotor}
    \caption{Attitude control testbed~\label{fig:hexrotor}}
\end{figure}
\vspace{-3mm}
A hexrotor unmanned aerial vehicle (UAV), as seen in~\cref{fig:hexrotor}, has been developed at the Flight Dynamics and Controls Laboratory (FDCL) at the George Washington University~\cite{kaufman2014}.
The UAV is composed of three pairs of counter-rotating propellers. 
Typical UAVs are composed of four or more co-planar propellers.
As a result, these systems are underactuated and unable to impart a force along every degree of freedom.
For example, quadrotor UAVs are unable to translate laterally without first conducting a rotation.
Conversely, the propeller pairs of the hexrotor are angled relative to one another to allow for a fully actuated rigid body.
This allows the hexrotor to impart a force in any direction and a moment about any axis. 
\begin{figure}[htbp]
    \centering 
    \subcaptionbox{ {Configuration error \( \Psi \)} \label{fig:Psi_tv} }{\input{pgf/Psi_timevarying.pgf}}~
    \subcaptionbox{ {Angle to constraint } \label{fig:con_angle_tv} }{\input{pgf/ang_con_timevarying.pgf}}~
    \subcaptionbox{ {Disturbance estimate \(\bar\Delta\) components} \label{fig:Delta_tv} }{\input{pgf/dist_timevarying.pgf}}
    \caption{Time-varying external disturbance simulation}
    \label{fig:tv} 
\end{figure}
\begin{figure}[t]
    \centering 
    \subcaptionbox{ {Attitude error vector \(e_R\)} components\label{fig:eR_exp} }{\input{pgf/eR_exp.pgf}}~
    \subcaptionbox{ {Configuration error \( \Psi \)} \label{fig:Psi_exp} }{\input{pgf/Psi_exp.pgf}}~
    \subcaptionbox{ {Control input \( u\) } components\label{fig:u_exp} }{\input{pgf/u_exp.pgf}}\\
    \subcaptionbox{ {Attitude Trajectory } \label{fig:traj_exp} }{\includegraphics[width=0.3\columnwidth]{traj_exp} }~
    \subcaptionbox{ {Angle to constraint } \label{fig:con_angle_exp} }{\input{pgf/ang_con_exp.pgf}}
    \caption{Constrained Attitude stabilization experiment}
    \label{fig:exp} 
\end{figure}
Attitude information is measured by a combination of both on and off board sensor systems.
The VectorNav VN-100 is a rugged, miniature high-performance inertial measurement unit which provides high frequency angular velocity measurements.
A Vicon motion capture system is installed within the test environment and used to provide high accuracy attitude measurements. 
A series of reflective markers are placed on the hexrotor and their relative position is captured by a series of infrared optical cameras. 
Assuming a fixed rigid body, the Vicon system is able to derive the attitude of the hexrotor and transmit this data to the processor onboard the hexrotor.
The control input is computed on-board, using the full state measurement, and implemented at approximately \SI{100}{\hertz}.
In order to constrain the motion, allowing us to test only the attitude dynamics, we attach the hexrotor to a spherical joint.
Since, the center of rotation is below the center of gravity of the hexrotor there is a destabilizing gravitational moment.
The resulting attitude dynamics are similar to an inverted pendulum model.
We augment the control input in~\cref{eqn:adaptive_control} with an additional term to negate the effect of the gravitational moment.

@@

\bibliography{BibMaster,library}
\bibliographystyle{IEEEtran}
